Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-09-30 draftfinalreport_v4.Rnw
New B/C analysis with text changes
Text changes are to include the new analysis range in the abstract, the
table, and the changes in the dhbcensus code.

vacc_dhb <- read.csv("tables/dhb_vacc.csv")
totals   <- vacc_dhb[nrow(vacc_dhb), ,drop=FALSE]
vacc_reqd_percent    <- totals$PC * 100
vacc_reqd_dhb_min    <- min(vacc_dhb$PC) * 100
vacc_reqd_dhb_max    <- max(vacc_dhb$PC) * 100
vacc_reqd_auckland   <- format(min(floor((vacc_dhb %>% filter(DHB == "Auckland" | DHB == "CountiesManakau" | DHB == "Waitemata") %>% select(Vaccination)) / 1000)) * 1000, big.mark=",")
vacc_reqd_count      <- format(totals$Vaccination, big.mark=",")
naive_percent        <- round(totals$Naive / totals$Population * 100)
break_even_vacc_cost <- format(74, big.mark=",") # vaccine costs such that benefits == cost (from .xlsx)
break_even_vacc_cost50 <- sprintf("%0.2f", 35.70) # vaccine costs such that benefits == cost (from .xlsx)
@

@@

\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is beneficial financially (B/C $>$1), with vaccination-related costs required to exceed approximately \$\Sexpr{break_even_vacc_cost} per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs together.

@@

change<-read.csv("change.csv",header=T)
costtable1<-read.csv("costtable1.csv",header=T)
costtable2<-read.csv("costtable2v2.csv",header=T)
costtable3<-read.csv("costtable3.csv",header=T)
costtable4<-read.csv("costtable4v2.csv",header=T)
costtable5<-read.csv("costtable5v2.csv",header=T)
costtable20<-read.csv("tables/cost_benefit_20.csv",header=T)
costtable50<-read.csv("tables/cost_benefit_50.csv",header=T)
costtable74<-read.csv("tables/cost_benefit_74.csv",header=T)
costtable20_no<-read.csv("tables/cost_benefit_no_20.csv",header=T)
costtable50_no<-read.csv("tables/cost_benefit_no_50.csv",header=T)
costtable_50pc<-read.csv("tables/cost_benefit_50_percent_vacc_35.csv",header=T)

@@

@@

The benefit--cost results for the scenario where the predicted number of measles cases based on epidemiological theory are expected to be prevented due to increased vaccination, in the absence of additional public measures, such as contact exclusion, are in \autoref{table:cost20_no} and \autoref{table:cost50_no}, for two different vaccine costs. The results in the two tables show the benefits of immunisation are always substantially greater than the costs of the increased supplementary immunisation (\autoref{table:cost20} and \autoref{table:cost50}).

@@

Our estimates of the costs of the measles outbreaks in New Zealand suggest that measles management costs per case in New Zealand is high. We used the mean values per case in the absence of alternative data and used all available data. The exact costs will vary, however, our benefit--cost analyses suggest that in all cases catch up immunisation schemes will be financially beneficial for New Zealand (see \autoref{table:cost20} and \autoref{table:cost50}). Additional vaccination beyond the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population we estimate are required to be vaccinated for $R_0 < 1$ decrease the cost effectiveness of an immunisation program. This is because large outbreaks are avoided once $R_0 < 1$. In the scenario we modelled with 50\% of the currently na\"{\i}ve population immunised, costs exceed benefits at \$\Sexpr{break_even_vacc_cost50} per immunised person \autoref{table:cost50pc}. There may be other benefits we have not accounted for in our analyses for additional vaccination. These additional benefits may include reducing the possibility of spatial heterogeneity and future benefits beyond the ten year period we modelled.

>>>>

vacc_dhb <- read.csv("tables/dhb_vacc.csv")
totals   <- vacc_dhb[nrow(vacc_dhb), ,drop=FALSE]
vacc_reqd_percent    <- totals$PC * 100
vacc_reqd_dhb_min    <- min(vacc_dhb$PC) * 100
vacc_reqd_dhb_max    <- max(vacc_dhb$PC) * 100
vacc_reqd_auckland   <- format(min(floor((vacc_dhb %>% filter(DHB == "Auckland" | DHB == "CountiesManakau" | DHB == "Waitemata") %>% select(Vaccination)) / 1000)) * 1000, big.mark=",")
vacc_reqd_count      <- format(totals$Vaccination, big.mark=",")
naive_percent        <- round(totals$Naive / totals$Population * 100)
break_even_vacc_cost <- format(74, big.mark=",") # vaccine costs such that benefits == cost (from .xlsx)
break_even_vacc_cost50 <- sprintf("%0.2f", 35.70) # vaccine costs such that benefits == cost (from .xlsx)
break_even_vacc_cost_no <- format(1877, big.mark=",")  # vaccine costs such that benefits == cost (from .xlsx)
@

@@

\item The benefit--cost (B/C) ratio analyses suggest additional vaccination is beneficial financially (B/C $>$1). Vaccination-related costs are required to exceed approximately \$\Sexpr{break_even_vacc_cost} to \$\Sexpr{break_even_vacc_cost_no} per person currently na\"{\i}ve immunized through vaccination before the costs exceed the benefits for all DHBs together, depending on the scenario modelled. In the former case, control measures, their costs and measles outbreak sizes are assumed to remain the same as 2009 onward, whereas in the latter example no control measures are implemented and the outbreak size is based on epidemiological theory in the absence of these control measures.

@@

change<-read.csv("change.csv",header=T)
costtable1<-read.csv("costtable1.csv",header=T)
costtable2<-read.csv("costtable2v2.csv",header=T)
costtable3<-read.csv("costtable3.csv",header=T)
costtable4<-read.csv("costtable4v2.csv",header=T)
costtable5<-read.csv("costtable5v2.csv",header=T)
costtable20<-read.csv("tables/cost_benefit_20.csv",header=T)
costtable50<-read.csv("tables/cost_benefit_50.csv",header=T)
costtable74<-read.csv("tables/cost_benefit_74.csv",header=T)
costtable20_no<-read.csv("tables/cost_benefit_no_20.csv",header=T)
costtable50_no<-read.csv("tables/cost_benefit_no_50.csv",header=T)
costtable1877_no<-read.csv("tables/cost_benefit_no_1877.csv",header=T)
costtable_50pc<-read.csv("tables/cost_benefit_50_percent_vacc_35.csv",header=T)

@@

\begin{sidewaystable}\footnotesize
<<results=tex,echo=FALSE>>=
colnames(costtable1877_no)<-c("DHB", "Vacc", "Vacc costs", "Wages saved", "Manage saved", "Hosp saved", "Contacts","Costs saved", "Outbreak", "OB costs", "B/C","NPV")
costtable1877_no[,c(2:10,12)] <- round(costtable1877_no[,c(2:10,12)])
costtable1877_no[,11] <- round(costtable1877_no[,11],2)
latex(costtable1877_no, file="", table.env=FALSE,rowname=NULL)
@
\caption{Benefit--cost analyses with \$\Sexpr{break_even_vacc_cost_no} dollars per immunised person assuming the cases prevented through additional vaccination each year are equal those based on epidemiological theory with no measles management costs (see \autoref{table:attack}). \textit{Vacc} is numbers to immunise (see \autoref{table:attack}); \textit{Vacc costs} is cost for the catch up immunisation programme; \textit{Wages saved} is wages of care givers and cases saved (\$839 per case); \textit{Contacts} is costs saved through contact exclusion (\$0 per case)\footnote{Here we assume no exclusions}; \textit{Manage saved} is management costs saved (\$0 per case), plus \$20 GP costs\footnote{Here we assume no public health management costs per measles case except GP visits}; \textit{Hosp saved} is the hospitalisation costs saved (\$1,877 per case); \textit{Costs save}  is the discounted costs saved; \textit{Outbreak}  is the predicted median outbreak size despite $R_v < 1$ due to measles importation from 1000 simulations\footnote{Note the median outbreak sizes are what is expected, while the mean takes into account rare, but large outbreaks, see \autoref{fig:sim} and \autoref{fig:sim1}.}; \textit{OB costs}  is costs expected due to continued measles importations based on 10 introductions of measles, one per year, but costs discounted on the same discounted rate; \textit{B/C} is the benefit--cost ratio; \textit{NPV} is the net present value, defined as the sum of the present values of costs and benefits of the vaccination programme.}
\label{table:cost1877_no}
\end{sidewaystable}

@@

The benefit--cost results for the scenario where the predicted number of measles cases based on epidemiological theory are expected to be prevented due to increased vaccination, in the absence of additional public measures, such as contact exclusion, are in \autoref{table:cost20_no} and \autoref{table:cost50_no}, for two different vaccine costs. The results in the two tables show the benefits of immunisation are always substantially greater than the costs of the increased supplementary immunisation (\autoref{table:cost20} and \autoref{table:cost50}). The cost of immunisation per person under this scenario, without public health measures and assuming homogeneous population mixing, was required to reach \$\Sexpr{break_even_vacc_cost_no} before the costs of additional immunisation exceeded the benefits (\autoref{table:cost1877_no}).

@@

Our estimates of the costs of the measles outbreaks in New Zealand suggest that measles management costs per case in New Zealand is high. We used the mean values per case in the absence of alternative data and used all available data. The exact costs will vary, however, our benefit--cost analyses suggest that in all cases catch up immunisation schemes will be financially beneficial for New Zealand (see \autoref{table:cost20} and \autoref{table:cost50}). Additional vaccination beyond the \Sexpr{vacc_reqd_percent}\% of currently na\"{\i}ve population we estimate are required to be vaccinated for $R_0 < 1$ decrease the cost effectiveness of an immunisation program. This is because large outbreaks are avoided once $R_0 < 1$. In the scenario we modelled with 50\% of the currently na\"{\i}ve population immunised, costs exceed benefits at \$\Sexpr{break_even_vacc_cost50} per immunised person \autoref{table:cost50pc}. There may be other benefits we have not accounted for in our analyses for additional vaccination. These additional benefits may include reducing the possibility of spatial heterogeneity and future benefits beyond the ten year period we modelled. Our analyses do suggest that in the absence of control measures the benefits of vaccination are substantial if measles were to spread through the na\"{\i}ve population as predicted by epidemiological theory (\autoref{table:cost20_no}, \autoref{table:cost50_no}, and \autoref{table:cost1877_no}).
